Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-07-28 paper.tex
updated the supplement to have the latest parsing results
marginally improved the drawing accuracy plot

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Table~\ref{drawingCommandTable}.  We predict the drawing
command token-by-token, conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. For
example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate.  See supplement for the full details of the
architecture, which we implemented in
Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}
The distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta \left[t_k | a_\theta \left(f_\theta(I,\text{render}(T)) | \{t_j\}_{j = 1}^{k - 1}\right), \{t_j\}_{j = 1}^{k - 1}\right]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network), and $a_\theta(\cdot|\cdot)$ is an attention mechanism. The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, the subscripts
on $T$ index drawing commands within the trace, and the \verb|STOP|
token is emitted by the network to signal that the trace
explains the image.
%% ~\remark{Make explicit that a $T_n$ in Equation 2
%%   is a concise way of referring to a sequence of tokens from Equation
%%   1?}

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes
and maximizing
(\ref{objective}) with respect to $\theta$ by gradient ascent.
Training does not require backpropagation across the entire sequence of drawing commands:
drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.
In a sense, this model is like an autoregressive variant of AIR~\cite{eslami1603attend} without attention.
We trained the network on $10^5$ scenes, which takes a little less than a day on an Nvidia TitanX GPU.

%% \remark{I like that you make this connection, but it could be made more precisely. Specifically, (1) the architecture isn't \emph{really} recurrent (it uses no hidden state cells), so it'd be good to use a different term or drop this part of the point: (2) training of recurrent nets is also typically fully-supervised (Most RNNs lack latent variables per timestep)---if you're thinking about AIR specifically, maybe just say that, and (3) it's like an autogressive AIR \emph{without attention}.}
%% \remark{Something related to this that's also cool to point out: training this model doesn't require backpropagation across the entire sequence of drawing commands (drawing to the canvas `blocks' the gradients, effectively offloading memory to an external (visual) store, so in principle it might be scalable to much longer sequences.}

%% When we
%% have the generative model (the rendering function) and treat the trace
%% as fully observed, we need not solve an unsupervised or reinforcement
%% learning problem.
\tikzset{>=latex}
\begin{figure}
  \begin{tikzpicture}
  \node[draw,blue,ultra thick,anchor = west,inner sep=0pt,label=below:Target image: $I$](observation) at (0,-1) {\includegraphics[width = 2cm]{figures/expert-18.png}};
  \node[draw,blue,thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 2cm,label=below:Canvas: render$(T)$] (canvas) at (0,-4) {};
  \draw[lightgray,ultra thin,step = 0.125] ([xshift = 0.5,yshift = 0.5]canvas.south west) grid ([xshift = -0.5,yshift = -0.5]canvas.north east);
  % draw partial image on canvas
  \draw (0.375cm,-3.25cm) circle (0.125cm);
  \draw (0.625cm,-3.625) circle (0.125cm);
  \draw (0.75cm,-3.375) circle (0.125cm);

  \node[draw,ultra thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 3cm] (CNN) at (4.5,-2.5) {CNN};
  \node[inner sep = 0pt](tensorProduct) at ([xshift = -1.5cm]CNN.west) {$\bigotimes$};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the execution trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\cite{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[h]
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes execution trace inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\caption{The deep network in (\ref{architecture}) predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

This network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathcal{R}$ uses the distance between two images as a proxy for a
likelihood.
%
Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We also compare with a natural baseline that models the problem as a kind of image captioning (``LSTM'' in Figure~\ref{syntheticResults}).
Given the target image, this baseline produces the program trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

\begin{figure}\centering
  \begin{minipage}{6cm}
    \begin{minipage}[t]{2.9cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-58-7.png}}\end{minipage}
    \begin{minipage}[t]{2.9cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-33-5.png}}\end{minipage}
    \caption{Network is trained to infer execution traces for randomly generated scenes like the two shown above. See supplement for details of the training data generation.}\label{trainingData}
  \end{minipage}\hfill%
  \begin{minipage}{7.5cm}\centering
    \includegraphics[width = 7cm]{figures/editDistanceRecurrent.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Using the model to parse latex output. The model is trained on diagrams with up to 12 objects. As shown above it generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
    \end{minipage}

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}; see supplement for details).
Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{equation}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{equation}
Intuitively,
$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
the trace we want and the trace we have so far.  Pixel-wise distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details. So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn carefully but not perfectly.
Our model assumes that objects are snapped to a $16\times 16$ grid, so
we made the drawings on graph paper.
\begin{figure}\centering
  \begin{minipage}{0.35\textwidth}
  \begin{minipage}[t]{\textwidth}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{0.5\textwidth}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{0.5\textwidth}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
  \end{minipage}\hfill
  \begin{minipage}{0.6\textwidth} 
  \includegraphics[width = 7.5cm]{figures/drawingAccuracy.png}
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,10,100)?
    Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 10 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace.}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-10.png}
    \includegraphics[width = 2cm]{figures/10-parse.png}    
    \subcaption{Noisy input}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-1.png}
    \includegraphics[width = 2cm]{figures/expert-1-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}     
    \subcaption{Failing on a Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}. Stochastic search (SMC) can help correct for these errors, as can the program synthesizer (Section~\ref{synthesizerHelpsParsing})}\label{lotsOfHandDrawings}%~\remark{Showing `failure cases' (the last three above) out-of-context seems not great. Perhaps remind people in the caption that stochastic search can help correct some of these issues?}
\end{figure}
For each drawing we annotated a ground truth trace and asked the neurally guided SMC sampler
to produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes, % on hand drawings,
we envision it working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

@@

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see~\cite{solar2008program}.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
See Figure~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.

@@

\begin{minipage}{\exampleTraceSize}\begin{verbatim}
Circle(center = (2,8)
Rectangle(6,9, 7,10)
Circle(center = (8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
Circle(center = (2,14)
Circle(center = (8,14)
Circle(center = (5,11)
Rectangle(3,12, 4,13)
\end{verbatim}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{verbatim}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{verbatim}
\end{minipage}\\\bottomrule
  \end{tabular}
  \caption{Example drawings (left), their ground truth traces (middle), and programs synthesized from these traces (right). Compared to the trace the programs are more compressive and automatically group together related drawing commands. are Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure, and affine transformation in the top figure.}\label{exampleSynthesisResults}
  \end{figure}

@@

\section{Conclusion}

@@

There are many directions for future work. In the parsing phase, the proposal network currently samples positional variables on a discrete grid. More general types of drawings could be supported by instead sampling from continuous distributions, e.g. using Mixture Density Networks~\cite{MDN}. The proposal network also currently handles only a very small subset of \LaTeX~drawing commands, though there is no reason that it could not be extended to handle more with a higher-capacity network. Exploring more sophisticated network architectures, including ones that utilize attention~\cite{mnih2014recurrent}, could help correct some of the errors the network makes.
In the synthesis phase, a more expressive DSL---including subroutines, recursion, and symmetry groups beyond reflections---would allow the system to effectively model a wider variety of graphical phenomena. The synthesizer itself could also be the subject of future work: the system currently uses the general-purpose Sketch synthesizer, which can take minutes to hours to run, whereas program synthesizers which are custom-built for special problem domains can run much faster or even interactively~\cite{le2014flashextract}.

>>>>

Figure~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Table~\ref{drawingCommandTable}.  We predict the drawing
command token-by-token, conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. For
example, the network first decides to emit the \verb|circle| token
conditioned on the image features, then it emits the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicts the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. For our main experiments we constrain
coordinates to lie on a discrete $16\times 16$ grid, and later explain
how to remove this restriction.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\cite{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

\tikzset{>=latex}
\begin{figure}[h]
  \begin{tikzpicture}
  \node[draw,blue,ultra thick,anchor = west,inner sep=0pt,label=below:Target image: $I$](observation) at (0,-1) {\includegraphics[width = 2cm]{figures/expert-18.png}};
  \node[draw,blue,thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 2cm,label=below:Canvas: render$(T)$] (canvas) at (0,-4) {};
  \draw[lightgray,ultra thin,step = 0.125] ([xshift = 0.5,yshift = 0.5]canvas.south west) grid ([xshift = -0.5,yshift = -0.5]canvas.north east);
  % draw partial image on canvas
  \draw (0.375cm,-3.25cm) circle (0.125cm);
  \draw (0.625cm,-3.625) circle (0.125cm);
  \draw (0.75cm,-3.375) circle (0.125cm);

  \node[draw,ultra thick,anchor = west,inner sep=0pt,minimum width = 2cm,minimum height = 3cm] (CNN) at (4.5,-2.5) {CNN};
  \node[inner sep = 0pt](tensorProduct) at ([xshift = -1.5cm]CNN.west) {$\bigoplus$};

@@

%  \node at (canvas.x,canvas.y) {Canvas};
\end{tikzpicture}
\caption{Our neural architecture for inferring the execution trace of a graphics program from its output. \textcolor{blue}{Blue}: network inputs. Black: network operations. \textcolor{red}{Red}: samples from a multinomial. \texttt{Typewriter font}: network outputs. Renders snapped to a $16\times 16$ grid, illustrated in \textcolor{gray}{gray}. STN (spatial transformer network) is a differentiable attention mechanism~\cite{jaderberg2015spatial}.}  \label{architecture}
\end{figure}
\begin{table}[h]
\begin{tabular}{ll}\toprule
  \begin{tabular}{l}
    \verb|circle|$(x,y)$
  \end{tabular}& \begin{tabular}{l}
    Circle at $(x,y)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|rectangle|$(x_1,y_1,x_2,y_2)$
  \end{tabular}&\begin{tabular}{l}
    Rectangle with corners at $(x_1,y_1)$ \& $(x_2,y_2)$
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|line|$(x_1,y_1,x_2,y_2,$\\
    \hspace{1cm}$\text{arrow}\in\{0,1\},\text{dashed}\in\{0,1\})$
  \end{tabular}&\begin{tabular}{l}
    Line from $(x_1,y_1)$ to  $(x_2,y_2)$,\\\hspace{1cm}optionally with an arrow and/or dashed
    \end{tabular}\\
  \begin{tabular}{l}
    \verb|STOP|
  \end{tabular}&\begin{tabular}{l}
    Finishes execution trace inference
    \end{tabular}
\\  \bottomrule
\end{tabular}
\caption{The deep network in Fig.~\ref{architecture} predicts drawing commands, shown above.}
\label{drawingCommandTable}
\end{table}

For the model in Fig.~\ref{architecture}, the distribution over the next drawing command factorizes as:
\begin{equation}
  \probability_\theta [t_1t_2\cdots t_K | I,T] = \prod_{k = 1}^K \probability_\theta \left[t_k | a_\theta \left(f_\theta(I,\text{render}(T)) | \{t_j\}_{j = 1}^{k - 1}\right), \{t_j\}_{j = 1}^{k - 1}\right]
\end{equation}
where $t_1t_2\cdots t_K$ are the tokens in the drawing command, $I$ is
the target image, $T$ is an execution trace, $\theta$ are the
parameters of the neural network, $f_\theta(\cdot,\cdot)$ is the
image feature extractor (convolutional network), and $a_\theta(\cdot|\cdot)$ is an attention mechanism. The distribution over
execution traces factorizes as:
\begin{equation}
  \probability_\theta [T|I] = \prod_{n = 1}^{|T|} \probability_\theta [T_n | I,T_{1:(n-1)}]\times\probability_\theta [\verb|STOP| | I,T]\label{objective}
\end{equation}
where $|T|$ is the length of execution trace $T$, the subscripts
on $T$ index drawing commands within the trace, and the \verb|STOP|
token is emitted by the network to signal that the trace
explains the image.

We train the network by sampling execution traces $T$ and target
images $I$ for randomly generated scenes
and maximizing
Eq.~\ref{objective} with respect to $\theta$ by gradient ascent.
We trained the network on $10^5$ scenes, which takes a little less than a day on an Nvidia TitanX GPU.

Training does not require backpropagation across the entire sequence
of drawing commands: drawing to the canvas `blocks' the gradients,
effectively offloading memory to an external visual store.  In a
sense, this model is like an autoregressive variant of
AIR~\cite{eslami1603attend}, but critically, \emph{without a learned
  recurrent memory state}, e.g. an RNN.  We can do without an RNN
because we have access to ground truth (image, trace) pairs.  This
allows handling scenes with a very large number of objects without
having to worry about the conditioning of gradients as they propagate
over long sequences, and also makes training more straightforward (it
is just maximum likelihood estimation of the model parameters).

Our network suffices to ``derender'' synthetic images like those shown in
Figure~\ref{trainingData}.  We can perform a beam search decoding to
recover what the network thinks is the most likely execution trace for
images like these, recovering traces maximizing $\probability_\theta
[T|I]$. But, if the network makes a mistake (predicts an incorrect
line of code), it has no way of recovering from the error.  In order
to derender an image with $n$ objects, it must correctly predict $n$
drawing commands -- so its probability of success will decrease
exponentially in $n$, assuming it has any nonzero chance of making a
mistake.  For added robustness as $n$ becomes large, we treat the
neural network outputs as proposals for a Sequential Monte Carlo (SMC) sampling scheme~\cite{SMCBook}.  For
the SMC sampler, we use pixel-wise distance as a surrogate for a
likelihood function. The SMC sampler is designed to produce samples
from the distribution $\propto L(I|\text{render}(T))
\probability_\theta[T|I]$, where $L(\cdot | \cdot):\text{image}^2\to
\mathbb{R}$ uses the distance between two images as a proxy for a
likelihood. Unconventionally, the target distribution of the SMC sampler
includes the likelihood under the proposal distribution.
Intuitively, both the proposal
distribution and the distance function offer complementary signals for
whether a drawing command is correct. Empirically, accuracy is much
better when we combine these signals.

Figure~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We also compare with a natural baseline that models the problem as a kind of image captioning (``LSTM'' in Figure~\ref{syntheticResults}).
Given the target image, this baseline produces the program trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

\begin{figure}[H]\centering
  \begin{minipage}{6cm}
    \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-58-7.png}}\end{minipage}%
    \begin{minipage}[t]{3cm}\fbox{\includegraphics[width=2.75cm]{figures/randomScene-33-5.png}}\end{minipage}
    \caption{Network is trained to infer execution traces for randomly generated scenes like the two shown above. See supplement for details of the training data generation.}\label{trainingData}
  \end{minipage}\hfill%
  \begin{minipage}{7.5cm}\centering
    \includegraphics[width = 7cm]{figures/editDistanceRecurrent.png}
%    \includegraphics[width = 4cm]{figures/accuracy.png}
      \caption{Parsing latex output after training on diagrams with $\leq 12$ objects. It generalizes to scenes with many more objects. Neither the stochastic search nor the neural network are sufficient on their own. \# particles varies by model: we compare the models \emph{with equal runtime} ($\approx 1$ sec/object)}\label{syntheticResults}
    \end{minipage}

@@

\subsection{Generalizing to hand drawings}
A practical application of our neural network is the automatic conversion of hand drawings into a subset of \LaTeX.
 We train the model
to generalize to hand drawings by introducing noise into the
renderings of the training target images.
We designed this noise process to introduce the kinds of variations found in hand drawings (Figure~\ref{handDrawingExamples}; see supplement for details).

\begin{wrapfigure}{R}{6cm}
\centering  \begin{minipage}[t]{2.5cm}\centering\includegraphics[width = 2cm]{figures/expert-60-reduced.png}
    \subcaption{}
  \end{minipage}\\
   \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-groundTruth-reduced.png}
    \subcaption{}
  \end{minipage}%
  \begin{minipage}[t]{2.5cm}\includegraphics[width = 2cm]{figures/60-1-reduced.png}
    \subcaption{}
  \end{minipage}%
    \caption{(a): a hand drawing. (b): Rendering of the trace our model infers for (a). We can generalize to hand drawings like these because we train the model on images corrupted by a noise process designed to resemble the kind of noise introduced by hand drawings - see (c) for a noisy rendering of (b).}\label{handDrawingExamples}
\end{wrapfigure}

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
\begin{multline}
-\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx\\ |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
\end{multline}
Intuitively,
$L_{\text{learned}}(\cdot|\cdot)$ approximates the distance between
the trace we want and the trace we have so far.  Pixel-wise distance
metrics are sensitive to the details of how
arrows, dashes, and corners are drawn -- but we wish to be invariant
to these details. So, we learn a distance metric over images that
approximates the distance metric in the search space over traces.

We drew 100 figures by hand; see figure~\ref{lotsOfHandDrawings}.
These were drawn carefully but not perfectly.
Because our model assumes that objects are snapped to a $16\times 16$ grid, 
we made the drawings on graph paper.

For each drawing we annotated a ground truth trace and asked the neurally guided SMC sampler
to produce many candidate traces for each drawing.
For 63\% of the drawings, the Top-1 most likely sample exactly matches the
ground truth; with more samples, the model finds traces
that are closer to the ground truth annotation (Fig.~\ref{drawingIntersectionOverUnion}).
Because our current model sometimes makes mistakes on hand drawings,
we envision it working as follows:
a user sketches a diagram,
and the system responds by proposing a few candidate interpretations.
The user could then select the one closest to their intention and edit it if necessary.

\begin{figure}[H]
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-73.png}
    \includegraphics[width = 2cm]{figures/73-parse.png}     
    \subcaption{A recurrent network architecture}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-21.png}
    \includegraphics[width = 2cm]{figures/21-parse.png}    
    \subcaption{A graphical model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-77.png}
    \includegraphics[width = 2cm]{figures/77-parse.png}    
    \subcaption{A figure from a deep learning textbook}
    \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-1.png}
    \includegraphics[width = 2cm]{figures/expert-1-parse.png}    
    \subcaption{Near miss}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-38.png}
    \includegraphics[width = 2cm]{figures/38-parse.png}     
    \subcaption{Ising model}
  \end{minipage}
  \begin{minipage}[t]{2.25cm}
    \includegraphics[width = 2cm]{figures/expert-34.png}
    \includegraphics[width = 2cm]{figures/34-parse.png}     
    \subcaption{Illusory contours (note: no SMC)}
    \end{minipage}
  \caption{Example drawings above model outputs. See also Fig.~\ref{firstPageExamples}. Stochastic search (SMC) can help correct for these errors, as can the program synthesizer (Section~\ref{synthesizerHelpsParsing})}\label{lotsOfHandDrawings}
\end{figure}

\begin{figure}[h]\centering
  \begin{minipage}{0.8\textwidth} 
\centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1,10,100)?
    Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 10 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace.}\label{drawingIntersectionOverUnion}
    \end{minipage}
\end{figure}

@@

The constrained optimization problem in
Equation~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}. Describing Sketch's program synthesis
algorithm is beyond the scope of this paper; see~\cite{solar2008program}.  At a
high level, Sketch takes as input a space of programs, along with a
specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a
minimum-cost program satisfying the specification.  In exchange for
having no guarantee on the time required to find a minimum
cost solution, it comes with the guarantee that it will always find a
globally optimal program.
In practice, synthesis times vary from a few minutes to a few hours,
with 16\% of the diagrams timing out the synthesizer after 3 hours.
See Figure~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.

@@

\begin{minipage}{\exampleTraceSize}\begin{verbatim}
Circle(center = (2,8)
Rectangle(6,9, 7,10)
Circle(center = (8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
Circle(center = (2,14)
Circle(center = (8,14)
Circle(center = (5,11)
Rectangle(3,12, 4,13)
\end{verbatim}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{verbatim}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{verbatim}
\end{minipage}\\\bottomrule
  \end{tabular}
  \caption{Example drawings (left), their ground truth traces (middle), and programs synthesized from these traces (right). Compared to the trace the programs are more compressive and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure, and affine transformation in the top figure.}\label{exampleSynthesisResults}
  \end{figure}

@@

\section{Discussion}

\subsection{Future work}
There are many directions for future work. The proposal network  currently handles only a very small subset of \LaTeX~drawing commands, though there is no reason that it could not be extended to handle more with a higher-capacity network.
In the synthesis phase, a more expressive DSL---including subroutines, recursion, and symmetry groups beyond reflections---would allow the system to effectively model a wider variety of graphical phenomena. The synthesizer itself could also be the subject of future work: the system currently uses the general-purpose Sketch synthesizer, which can take minutes to hours to run, whereas program synthesizers which are custom-built for special problem domains can run much faster or even interactively~\cite{le2014flashextract}.

In the parsing phase, the proposal network currently samples positional variables on a discrete $16\times 16$ grid. More general types of drawings could be supported by instead sampling from continuous distributions, e.g. using Mixture Density Networks~\cite{MDN}.
Implementing a variant of our model that predicts continuous coordinates
using Mixture Density Networks is straightforward;
we have prototyped such a system and
show examples of it working in the supplement.
But, for our collection of hand drawings, discrete positional variables suffice.

\subsection{Contributions}

@@
